Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-13 eswc-2012/phdsymp.tex
methods in fed eval
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@95 43ef1190-90fd-4dd7-a769-d0121f02b80e
skip treatment
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@96 43ef1190-90fd-4dd7-a769-d0121f02b80e

In this case, we must consider the possibility that the measurements
are impacted by random noise. Sources of random noise may be the
overall system load, garbage collection triggers, use of randomized
algorithms, or the data of such a nature that it triggers worst-case
behaviour, etc. All this contributes to that we should consider the
performance metric, whatever metric is used, as a \emph{stochastic
  variable}.

@@

The problems addressed by existing benchmarks such as the one quoted
above are almost orthogonal to the problems considered by my proposed
project. While I have heard of some cases where a hypothesis test is
used to demonstrate that one implementation is better than another, it
is seemingly not common practice. Furthermore, I have not to date seen
any work towards using contemporary statistical methods to evaluate
the performance of software, but I shall admit that my literature
study have been limited to RDF/OWL databases, and practical benchmarks
of databases, file systems, scientific software, etc. Thus, the
existence of relevant references in the deeper computer science
literature is a key issue I would appreciate discussing at the
Symposium.

@@

@@

This part of the thesis seeks to use advances from statistical
experimental planning to improve accuracy and dependability of
benchmarking. Already in 1926, Ronald~Fischer noted that complex
experiments can be much more efficient than simple ones\footnote{Cited
  in http://en.wikipedia.org/wiki/Factorial\_experiment}, starting the
experimental design field. One of the simpler designs is ``fractional
factorial design'', in which several ``factors'' are studied by using
different ``treatments''. In terms of SPARQL execution, the SPARQL
engine is clearly a factor, but also for example the nestedness of
\texttt{OPTIONAL}s can be a factor, number of triples in a basic graph
pattern, etc, and the treatment is to vary these numbers to different
``levels''. The key to understanding why this can be efficient is that
these variations need not occur in the same experiment. Thus, many
combinations of factors can be studied by carefully designing queries
to cover different factors, and a formalism called ``resolution'' has
been developed to classify how well this has been achieved.

>>>>

In this case, we must consider the possibility that the measurements
are impacted by random noise. Sources of random noise may be the
overall system load, garbage collection triggers, use of randomized
algorithms, or that the data are of such a nature that it triggers
worst-case behaviour, etc. All this contributes to that we should
consider the performance metric, whatever metric is used, as a
\emph{stochastic variable}.

@@

The problems addressed by existing benchmarks such as the one cited
above are almost orthogonal to the problems considered by my proposed
project. While I have heard of some cases where a hypothesis test is
used to demonstrate that one implementation is better than another, it
is seemingly not common practice. Furthermore, I have not to date seen
any work towards using contemporary statistical methods to evaluate
the performance of software, but I shall admit that my literature
study have been limited to RDF/OWL databases, and practical benchmarks
of databases, file systems, scientific software, etc. Thus, the
existence of relevant references in the deeper computer science
literature is a key issue I would appreciate discussing at the
Symposium.

@@

The evaluation methodology for the SPARQL federation work of the
thesis will largely be covered by running the elaborate benchmark
designs of the thesis.

@@

This part of the thesis seeks to use advances from statistical
experimental planning to improve accuracy and dependability of
benchmarking. Already in 1926, Ronald~Fischer noted that complex
experiments can be much more efficient than simple ones\footnote{Cited
  in http://en.wikipedia.org/wiki/Factorial\_experiment}, starting the
experimental design field. One of the simpler designs is ``fractional
factorial design'', in which several ``factors'' are studied. In terms
of SPARQL execution, the SPARQL engine is clearly a factor, but also
for example the nestedness of \texttt{OPTIONAL}s can be a factor,
number of triples in a basic graph pattern, etc, and these numbers are
varied to different ``levels''. The key to understanding why this can
be efficient is that these variations need not occur in the same
experiment. Thus, many combinations of factors can be studied by
carefully designing queries to cover different factors, and a
formalism called ``resolution'' has been developed to classify how
well this has been achieved, partly answering the question of
evaluation methodology for this part of the thesis.
